### 2019-08-16 HRS_plotting.Rmd
manuscript changes, add some R scripts used in analysis and dataviz

---
title: "HRS plotting"
author: 
  - name: Alex Diaz-Papkovich
affiliation:
  - &cruk Quantitative Life Sciences, McGill University,  Montreal, Canada
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  html_notebook:
  df_print: paged
code_folding: show
toc: yes
toc_float: 
  collapsed: false
smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HRS admixture plots

teeeeeext

## Including Plots

```{r}
library(tidyverse)
library(ggplot2)
library(Ternary)
library(dplyr)
```

# Import the HRS auxiliary data and projection to use.
```{r}
# Auxiliary data
hrs_aux_data <- read.csv("/Volumes/Stockage/alex/hrs/aux/HRS_aux.csv")

# Projection
fname <- "HRS_1000G_NP1_UMAP_PC10_NC2_NN15_MD0.5_pca_1kgp_onto_hrs_umap_1kgp_onto_hrs_2018112221116"
proj_dir <- "/Volumes/Stockage/alex/gt-dimred/np1/projections"

hrs_1kgp_proj <- read.table(paste(proj_dir, fname, sep="/"))

# This is HRS+1KGP data; first 12,454 are HRS
hrs_proj <- hrs_1kgp_proj[1:12454,]

# Merge the tables
hrs_full_data <- cbind(hrs_aux_data, hrs_proj)

# RGB colours from admixture
hrs_full_data$rgb_vals <- rgb(hrs_full_data$ADMIX1*255,
                              hrs_full_data$ADMIX2*255,
                              hrs_full_data$ADMIX3*255,
                              maxColorValue = 255)

```

```{r}
# legends
require(RColorBrewer)
#pcol=brewer.pal(9, "Set1")

AncCol=c("#FF0000","#00FF00","#0000FF")

AncRGB=col2rgb(AncCol)/255
#pdf("UMAP_ternary_plot_legend.pdf",height=2.375,width=2.375,pointsize=10)
par(mar=c(2.1,2.1,0.1,0.1))
TernaryPlot(alab=expression(''%->% "African ancestry (%)" %->% ''),blab=expression(''%->%"European ancestry (%)" %->% ''),clab=expression(''%<-% "Asian/Native American ancestry (%)" %<-% ''),lab.cex=1,axis.cex = 0.8)
dx=0.01;
cnt=0
bk=0;
for(xi in seq(0,1,length=round(1/dx))){
  for(yi in seq(0,1,length=round(1/dx))){
    for(zi in seq(0,1,length=round(1/dx))){
      if(abs(1-(xi+yi+zi)) < 1e-8){
        ired = xi*AncRGB[1,1] + yi*AncRGB[1,2] + zi*AncRGB[1,3]
        igreen = xi*AncRGB[2,1] + yi*AncRGB[2,2] + zi*AncRGB[2,3]
        iblue = xi*AncRGB[3,1] + yi*AncRGB[3,2] + zi*AncRGB[3,3]
        AddToTernary(points,c(xi,yi,zi),col=rgb(ired,igreen,iblue,1),pch=20)
        cnt=cnt+1
      }
    }
  }
}

# legends
require(RColorBrewer)
#pcol=brewer.pal(9, "Set1")

AncCol=c("#FF0000","#0000FF","#00FF00")

AncRGB=col2rgb(AncCol)/255
#pdf("UMAP_ternary_plot_legend.pdf",height=2.375,width=2.375,pointsize=10)
par(mar=c(2.1,2.1,0.1,0.1))
TernaryPlot(alab=expression(''%->% "African ancestry (%)" %->% ''),blab=expression(''%->%"European ancestry (%)" %->% ''),clab=expression(''%<-% "Asian/Native American ancestry (%)" %<-% ''),lab.cex=1,axis.cex = 0.8)
dx=0.01;
cnt=0
bk=0;
for(xi in seq(0,1,length=round(1/dx))){
  for(yi in seq(0,1,length=round(1/dx))){
    for(zi in seq(0,1,length=round(1/dx))){
      if(abs(1-(xi+yi+zi)) < 1e-8){
        ired = xi*AncRGB[1,1] + yi*AncRGB[1,2] + zi*AncRGB[1,3]
        igreen = xi*AncRGB[2,1] + yi*AncRGB[2,2] + zi*AncRGB[2,3]
        iblue = xi*AncRGB[3,1] + yi*AncRGB[3,2] + zi*AncRGB[3,3]
        AddToTernary(points,c(xi,yi,zi),col=rgb(ired,igreen,iblue,1),pch=20)
        cnt=cnt+1
      }
    }
  }
}
```

```{r}
out_dir <- "/Volumes/Stockage/alex/hrs_1000G/images/for_export/alt_colours"
fname <- "HRS_1000G_NP1_UMAP_PC10_NC2_NN15_MD05_pca_1kgp_onto_hrs_umap_1kgp_onto_hrs_2018112221116_birth"

p_hrs_birth <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=BirthRegionName)) +
  geom_point(alpha=0.6, size=0.1) +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="bottom",
        legend.title=element_blank()) + 
  guides(colour = guide_legend(override.aes = list(size=4, alpha = 1)))

p_hrs_birth

ggsave(paste(out_dir, "/", fname, ".jpeg", sep=""), height=h, width=w, units=u)
```

```{r}
p_hrs <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=rgb(hrs_full_data$ADMIX1*255,
                                                       hrs_full_data$ADMIX2*255,
                                                       hrs_full_data$ADMIX3*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=0.1) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs
```

# Alternate colourings
```{r}

out_dir <- "/Volumes/Stockage/alex/hrs_1000G/images/for_export/alt_colours"
fname <- "HRS_1000G_NP1_UMAP_PC10_NC2_NN15_MD05_pca_1kgp_onto_hrs_umap_1kgp_onto_hrs_2018112221116_admix"

# dimensions
h <- 120.4
w <- 186.0

# units
u <- "mm"

p_hrs <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=rgb(hrs_full_data$ADMIX1*255,
                                                       hrs_full_data$ADMIX2*255,
                                                       hrs_full_data$ADMIX3*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=0.1) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "123.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=rgb(hrs_full_data$ADMIX1*255,
                                                       hrs_full_data$ADMIX3*255,
                                                       hrs_full_data$ADMIX2*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=0.1) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "132.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=rgb(hrs_full_data$ADMIX2*255,
                                                       hrs_full_data$ADMIX1*255,
                                                       hrs_full_data$ADMIX3*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=0.1) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "213.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=rgb(hrs_full_data$ADMIX2*255,
                                                       hrs_full_data$ADMIX3*255,
                                                       hrs_full_data$ADMIX1*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=0.1) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "231.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=rgb(hrs_full_data$ADMIX3*255,
                                                       hrs_full_data$ADMIX1*255,
                                                       hrs_full_data$ADMIX2*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=0.1) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "312.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_full_data, aes(x=V1, y=V2, col=rgb(hrs_full_data$ADMIX3*255,
                                                       hrs_full_data$ADMIX2*255,
                                                       hrs_full_data$ADMIX1*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=0.1) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "321.jpeg", sep=""), height=h, width=w, units=u)
```

# Same, but for Hispanic data
```{r}

# Auxiliary data
#hrs_aux_data <- read.csv("/Volumes/Stockage/alex/hrs/aux/HRS_aux.csv")

# Projection
fname <- "HRS_1000G_NP1_UMAP_PC7_NC2_NN15_MD0.5_pca_hrshisp_added1kgp_2018115153245"
proj_dir <- "/Volumes/Stockage/alex/gt-dimred/np1/projections"

hrs_hisp_1kgp_proj <- read.table(paste(proj_dir, fname, sep="/"))

# Only the first 12,454 are HRS.
# Note that we did the projection with all the HRS (so we can also project the other HRS onto Hisps)
hrs_hisp_1kgp_proj <- hrs_hisp_1kgp_proj[1:12454,]

# Merge the tables
hrs_hisp_data <- cbind(hrs_aux_data, hrs_hisp_1kgp_proj)

# Reduce to just Hispanic individuals
hrs_hisp_data <- subset(hrs_hisp_data, HispanicStatus=="Hispanic")

# RGB colours from admixture
hrs_hisp_data$rgb_vals <- rgb(hrs_hisp_data$ADMIX1*255,
                              hrs_hisp_data$ADMIX2*255,
                              hrs_hisp_data$ADMIX3*255,
                              maxColorValue = 255)

```

# Create beautiful plots!
```{r}
out_dir <- "/Volumes/Stockage/alex/hrs_1000G/images/for_export/alt_colours"
fname <- "HRS_1000G_NP1_UMAP_PC7_NC2_NN15_MD05_pca_hrshisp_added1kgp_2018115153245_admix"

# dimensions
h <- 154.6
w <- 169.2

# units
u <- "mm"

# point size
s <- 0.5

p_hrs <- ggplot(hrs_hisp_data, aes(x=V1, y=V2, col=rgb(hrs_hisp_data$ADMIX1*255,
                                                       hrs_hisp_data$ADMIX2*255,
                                                       hrs_hisp_data$ADMIX3*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=s) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "123_hisp.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_hisp_data, aes(x=V1, y=V2, col=rgb(hrs_hisp_data$ADMIX1*255,
                                                       hrs_hisp_data$ADMIX3*255,
                                                       hrs_hisp_data$ADMIX2*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=s) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "132_hisp.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_hisp_data, aes(x=V1, y=V2, col=rgb(hrs_hisp_data$ADMIX2*255,
                                                       hrs_hisp_data$ADMIX1*255,
                                                       hrs_hisp_data$ADMIX3*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=s) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "213_hisp.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_hisp_data, aes(x=V1, y=V2, col=rgb(hrs_hisp_data$ADMIX2*255,
                                                       hrs_hisp_data$ADMIX3*255,
                                                       hrs_hisp_data$ADMIX1*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=s) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "231_hisp.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_hisp_data, aes(x=V1, y=V2, col=rgb(hrs_hisp_data$ADMIX3*255,
                                                       hrs_hisp_data$ADMIX1*255,
                                                       hrs_hisp_data$ADMIX2*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=s) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "312_hisp.jpeg", sep=""), height=h, width=w, units=u)

p_hrs <- ggplot(hrs_hisp_data, aes(x=V1, y=V2, col=rgb(hrs_hisp_data$ADMIX3*255,
                                                       hrs_hisp_data$ADMIX2*255,
                                                       hrs_hisp_data$ADMIX1*255,
                                                       maxColorValue = 255))) +
  geom_point(alpha=0.6, size=s) +
  scale_color_identity() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

p_hrs

ggsave(paste(out_dir, "/", fname, "321_hisp.jpeg", sep=""), height=h, width=w, units=u)
```

```{r}
library('Ternary')
require(RColorBrewer)
#pcol=brewer.pal(9, "Set1")

AncCol=c("#4DAF4A","#E41A1C","#377EB8")

AncRGB=col2rgb(AncCol)/255
#pdf("UMAP_ternary_plot_legend.pdf",height=2.375,width=2.375,pointsize=10)
par(mar=c(2.1,2.1,0.1,0.1))
TernaryPlot(alab=expression(''%->% "African ancestry (%)" %->% ''),blab=expression(''%->%"European ancestry (%)" %->% ''),clab=expression(''%<-% "Native American ancestry (%)" %<-% ''),lab.cex=0.6,axis.cex = 0.5)
dx=0.01;
cnt=0
bk=0;
for(xi in seq(0,1,length=round(1/dx))){
  for(yi in seq(0,1,length=round(1/dx))){
    for(zi in seq(0,1,length=round(1/dx))){
      if(abs(1-(xi+yi+zi)) < 1e-8){
        ired = xi*AncRGB[1,1] + yi*AncRGB[1,2] + zi*AncRGB[1,3]
        igreen = xi*AncRGB[2,1] + yi*AncRGB[2,2] + zi*AncRGB[2,3]
        iblue = xi*AncRGB[3,1] + yi*AncRGB[3,2] + zi*AncRGB[3,3]
        AddToTernary(points,c(xi,yi,zi),col=rgb(ired,igreen,iblue,1),pch=20)
        cnt=cnt+1
      }
    }
  }
}
```

```{r}
tern_data <- select(hrs_full_data,ADMIX1*255,ADMIX2*255,ADMIX3*255)

p_tern <- ggtern() +
  geom_point(data=data.frame(x=255,y=255,z=255), aes(x,y,z))

p_tern
```