### 2019-08-16 UKBB_plotting.Rmd

---
title: "London alpha plotting"
author: 
  - name: Alex Diaz-Papkovich
affiliation:
  - &cruk Quantitative Life Sciences, McGill University,  Montreal, Canada
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  html_notebook:
  df_print: paged
code_folding: show
toc: yes
toc_float: 
  collapsed: false
smooth_scroll: false
---
## Plotting UKBB stuff
Python's plotting is getting on my nerves. I want to make a scatterplot where I specify each point's colour and alpha values.

# Import libraries
```{r}
library(ggplot2)
library(tidyverse)
```

# Import the data
```{r}
data_dir <- "/Volumes/Stockage/alex/ukbb_other/data/london_alphas.csv"

ukbb_plotting_data <- read.csv(data_dir)
ukbb_plotting_data <- ukbb_plotting_data[!(ukbb_plotting_data$eth_txt %in% c('British', 'Irish', 'Any other white background')),]

out_dir <- "/Volumes/Stockage/alex/ukbb_images/london_alpha"
fname <- "UKBB_UMAP_PC10_NN15_MD05_2018328174511_within50km"
 h <- 15
 w <- 15

p_scatter <- ggplot(ukbb_plotting_data, aes(x=x_coords, y=y_coords)) +
  geom_point(colour=ukbb_plotting_data$colour, alpha=ukbb_plotting_data$London_alpha, stroke=0) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())
p_scatter

ggsave(paste(out_dir, paste(fname,".jpeg",sep=""), sep="/"), device = "jpeg", height = h, width = w, units = "in")
```

# What about those outside of 50k from London?
```{r}
data_dir <- "/Volumes/Stockage/alex/ukbb_other/data/london_alphas.csv"

ukbb_plotting_data <- read.csv(data_dir)
ukbb_plotting_data <- ukbb_plotting_data[!(ukbb_plotting_data$eth_txt %in% c('British', 'Irish', 'Any other white background')),]

ukbb_plotting_data$Outside_London <- ifelse(ukbb_plotting_data$London_alpha==0,1,0)

out_dir <- "/Volumes/Stockage/alex/ukbb_images/london_alpha"
fname <- "UKBB_UMAP_PC10_NN15_MD05_2018328174511_outside_50km"
 h <- 15
 w <- 15

p_scatter <- ggplot(ukbb_plotting_data, aes(x=x_coords, y=y_coords)) +
  geom_point(colour=ukbb_plotting_data$colour, alpha=ukbb_plotting_data$Outside_London, stroke=0) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())
p_scatter

ggsave(paste(out_dir, paste(fname,".jpeg",sep=""), sep="/"), device = "jpeg", height = h, width = w, units = "in")
```